---
title: "Population Genetice Measures: F~st~"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Population Genetice Measures: F~st~}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we show an example of `evodeepo`'s functions for calculating population genetic measure, in this case F~st~. The main reason to use `evodeepo` for this is speed. Though this technically has nothing in particular to do with deep learning, which is the main point of `evodeepo`, the deep learning library on which `evodeepo` is based `torch`, is highly optimized and so by using it to calculate these measures we can achieve a massive speedup over calculations done in base R. Here we demonstrate the `evd_fst()` function and benchmark it against `StAMPP::stamppFst()` (though not they calculate different estimates of Fst, but we will talk more about that later on). Let's load the packages we will need.

```{r setup}
library(evodeepo)
library(dartR)
library(StAMPP) 
```

For this example we will use an dataset in the dartR package with SNP data on bandicoots. Later we will use some simulated data to test the method on a much larger number of SNPs. Let's try calculating F~st~ directly on this bandicoot `genlight` object. 

```{r}
bandicoot_fst <- evd_fst(bandicoot.gl)
```
